### 2025-01-07 README.Rmd
updating README

---
output: github_document
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
```

# HMMotion: Using tracking data to predict coverage strategies in American Football <img src="plots/bdb_logo.png" align="right" height=170>

This is the github repository for the corresponding submission to the NFL Big Data Bowl 2024 competition on Kaggle. The full notebook can be viewed [here](https://www.kaggle.com/code/rouvenmichels/hmmotion-using-tracking-data-to-predict-coverage).

## Quick Summary

<center>
<img src = "./animation/animation.gif" width = 500>
</center>

## Code Information